---
title: "HW1"
author: "Yui Kan Kong"
date: "2022-09-24"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Before anything else

Download the R package AER. This package contains a number of data sets.
Upload the data set "USInvest" into a data frame as a time series. You
will find the data set listed under Greene (2003).

```{r cars}

if(!require(dplyr)){install.packages('dplyr')}
if(!require(AER)){install.packages('AER')}
if(!require(gridExtra)){install.packages('gridExtra')}
data(USInvest)                                   # Load dataset
data=USInvest                        # Assign to a variable
data
print("Check is it time series as requested in the question")
is.ts(data)
print("To make life easies, trandorm to dataframe")
mydata.df <- data.frame(year=time(data),as.matrix(data))
mydata.df

```

a)  Add real GNP, real invest, and inflation to the data frame.
    Descibtion from orginal document An annual multiple time series from
    1968 to 1982 with 4 variables.

gnp Nominal gross national product, invest Nominal investment, price
Consumer price index, interest Interest rate (average yearly discount
rate at the New York Federal Reserve Bank). From internet RGNP =
(GNP/CPI) x 100. Real investment=(investment/CPI)x100
inflation=$(CPI_{t} + CPI_{t-1})/CPI_{t-1} \times 100$

```{r a}
mydata.df$rgnp=mydata.df$gnp/mydata.df$price*100
mydata.df$real.investment=mydata.df$invest/mydata.df$price*100
mydata.df$price_shifted <- c(NA, head(mydata.df$price, -1))
mydata.df$inflation=(mydata.df$price-mydata.df$price_shifted)/mydata.df$price_shifted  
mydata.df
```

b)  Calculate the mean and standard deviation for each of the series.

```{r}
AllTheMeans <- sapply(mydata.df,mean)
print("All mean other than inflation")
AllTheMeans 
print("inflation mean")
mean(mydata.df$inflation, na.rm=TRUE)
print("all other standard deviation")
AllThesd <- sapply(mydata.df,sd)
AllThesd
print("Inflation sd")
sd(mydata.df$inflation, na.rm=TRUE)

```

c)  Include growth rates for real GNP and real investment in the data
    frame.

```{r}
library(dplyr)
mydata.df = mydata.df %>%
  # first sort by year
  arrange(year) %>%
  mutate(Diff_year = year - lag(year),
         Diff_rgnp = rgnp - lag(rgnp),
         Diff_real.investment = real.investment - lag(real.investment),
         Growth_in_rgnp = (Diff_rgnp / Diff_year)/rgnp * 100,
         Growth_in_real.investment = (Diff_real.investment / Diff_year)/real.investment * 100) 

mydata.df
```

d)  Plot real GNP, real invest, and inflation as lines on separate
    charts on a single page. Be sure to put time on the horizontal axis.
    Include a legends and a title.

```{r}

library(ggplot2)

ggplot(data=mydata.df, aes(x=year,y=rgnp))+geom_line(aes(color = "Real GNP"))+ theme(legend.position="top")+ ggtitle("Real GNP")
```

```{r}
ggplot(data=mydata.df, aes(x=year,y=real.investment))+geom_line(aes(color = "Real Investment"))+ theme(legend.position="top")+ ggtitle("Real investment")
```

```{r}
ggplot(data=mydata.df, aes(x=year,y=inflation))+geom_line(aes(color = "Inflation"))+ theme(legend.position="top")+ ggtitle("Inflation")

```

e)  Plot the growth rates of real GNP and real investment on separate
    charts and include a horizontal line on each chart representing the
    mean rate of growth for the series.

The black line is the mean

```{r}
ggplot(data=mydata.df, aes(x=year,y=Growth_in_rgnp))+geom_line(aes(color = "Growth_in_rgnp"))+ theme(legend.position="top")+ ggtitle("Growth_in_rgnp")+geom_hline(yintercept=mean(mydata.df$Growth_in_rgnp, na.rm=TRUE))

```

```{r}
ggplot(data=mydata.df, aes(x=year,y=Growth_in_real.investment))+geom_line(aes(color = "Growth_in_real.investment"))+ ggtitle("Growth_in_real.investment")+geom_hline(yintercept=mean(mydata.df$Growth_in_real.investment, na.rm=TRUE),show.legend = "mean")+ theme(legend.position="top")

```
